Differential Testing Runtime Engines--Lessons Learned from the JavaScript Domain

### 2018-07-30 paper/paper.tex
pass on result
adjusting text

JavaScript (\js{}) is today one of the most popular programming
languages for the web~\cite{business-insider,stackify}. The interest
of the community for the language encourages constant improvements in
its specification and implementations (see
Section~\ref{sec:es6-design}). It is natural to expect that such
improvements will entail sensible changes in runtime
engines~\cite{kangax}; changes that could lead to errors. Even small
changes can have a large impact in practice. For example, in October
2014 a new attribute added to the Array datatype caused a bug that
resulted in MS Outlook web not working under
Chrome~\cite{array-bug-chromium-issue4247,array-bug-discussion}. To
sum, finding bugs on JS engines is an important problem.

Differential testing~\cite{Brumley-etal-ss07} (DT) is a popular
approach that has been applied in a variety of contexts to find bugs
in
software~\cite{Yang-etal-pldi11,Chen-etal-fse2015,Argyros-etla-ccs16,Chen-etal-pldi16,petsios-etal-sp2017,SivakornAPKJ17}. It
automates test generation in scenarios where multiple implementations
of a system exist. DT leverages the diversity across system's
implementations to detect anomalous behavior. For example, Mozilla
runs JS files--created with the grammar-based fuzzer
jsfunfuzz~\cite{jsfunfuzz}--against different configurations of their
SpiderMonkey engine~\cite{jsfunfuzz-mozilla-bug} to find discrepancies
across configurations. They use this technique since 2002 and have
been able to find over 270 bugs since
then~\cite{jsfunfuzz-at-mozilla}, some of which are security bugs.
Cross-engine differential testing, in contrast, checks correctness by
comparing the output produced by \emph{different} engines on a given
input~\cite{patra2016learning}. The availability of public
implementations of JS engines (see Table~\ref{tab:engines}) and test
suites makes this bug-finding approach appealing.

%% WHAT WE DID
This paper reports the results of a study we conducted to (i) assess
the reliability of existing JS engines and to (ii) evaluate the
importance of different factors when using DT. More precisely, we
analyzed the impact of two orthogonal factors in DT: test mining and
fuzzing. The first factor, test mining, considers the effect that
\js{} test files mined from the repositories of open-source engines
have in detecting functional bugs in other engines. The second factor,
fuzzing, considers the impact in bug detection of fuzzing these files.
%% WHAT WE FOUND
To sum, we found that \Fix{...}

@@

\section{EcmaScript}
\label{sec:es6-design}

ECMAScript (\es{}) emerged with the goal to standardize variants of
JavaScript, such as Netscape's JavaScript and Microsoft's JScript. The
\es{} specification is maintained by the TC39~\cite{tc39-github}
technical committee under the regulation of Ecma
International~\cite{es6-website}.  Every year, a new version of the
specification is delivered with new features and minor fixes. The
canonical spec today is
ES6~\cite{ecmas262-spec-repo,ecmas262-spec}. Features of these
versions and supporting engines can be found elsewhere~\cite{kangax}.

\subsection{Engines Design}

\Mar{what is that exactly? can we cite? is that only spec or actually
  code as most of java.* libraries.}  \Igor{seems like an API. They
  define the language behavior; grammar; values to static variables;
  the methods for each "class"(Object)...  These irregularities
  appears on engine implementation, maybe an wrong if statement and so
  on.} \Mar{Igor, avoid general discussion like $\leftarrow$this. look for
  explanation in some forum/tutorial and explain **when** you
  undertand the JS ecosystem design and can explain with confidence.}

\subsection{Implementation-Dependent Behavior}
\label{sec:imp-dep-behavior}

The \es{} spec uses the label ``implementation-dependent'' to indicate
cases where each engine developer can decide how to implement some
aspect of a functionality. In these cases, execution may differ from
engine to engine. One of the reasons implementation-dependent behavior
exists is to enable compiler optimizations. For example, the
\mcodeid{for-in} loop does not clearly specify the iteration
order~\cite{for-in-undefined} and different engines capitalize on
that.

@@

\subsection{Undefined Behavior}
\label{sec:undef-behavior}

Undefined behavior is not a feature of the C
language~\cite{c-undefined}.

\Mar{mention strict?}

\subsection{Unintentional Discrepancies}

@@

\subsection{Randomness and Undefined Behavior}

\Mar{answer from C. Holler. Think through this...}

1) The stack depth in JS is often close to the real (native) C stack depth, nevertheless observable by JS through recursion. This causes differential behavior even on the same engine when JITs are enabled or disabled, as the stack frame sizes vary. LangFuzz in particular produces a lot of recursions and hits these problems even more often than jsfunfuzz.

2) JS has some "undefined behavior", such as the enumeration order of object properties in certain cases. In our engine, we have an --enable-more-deterministic mode (build option) to solve some of these issues.

3) JS naturally has some observable differential behavior in global objects, e.g. Math.random, Date, etc. These can often be mocked to solve these issues, but fuzzers can be surprisingly effective to find more of these sources or even get the old objects back somehow.

@@

\subsection{Seed JS files\label{sec:seeds}}~
We selected test files from several public projects from GitHub. We
looked for self-contained tests from JS engine projects, initially
using the GitHub REST API~\cite{github-rest-api}. We noticed that some
of the tests we found depend on external libraries, which not all
engines we use (see Section~\ref{sec:methodology:engines}) support. We
decided to discard those. For example, many tests we found required a
Node.js runtime~\cite{node} for execution. We considered the offical
TC-39~\cite{tc39-github} conformance test suite for the ECMA262
specification~\cite{ecmas262-spec}. Furthermore, we considered test
suites of three of the four engines mentioned in
Section~\ref{sec:methodology:engines} and four other engines.
Table~\ref{tab:test-suites} shows the JS sources we considered for
each of the selected projects. Overall, we found a total of
\totfiles{} JS files. It is worth noting that some test suites use
distinct frameworks for testing and require external libraries to run
their testing suites.  For these projects, we needed to make small
changes in the testing infrastructure to be able to run the tests
uniformly across all engines. More precisely, we needed to mock
functions, related to the test harness, which are only available to
certain engines.

@@

Generational fuzzers are typically grammar-based. These fuzzers
generate a new file using the grammar of the language whose inputs
should be fuzzed. Intuitively, those fuzzers implement a traversal of
the production rules of the input grammar to create syntax trees,
which are then pretty-printed. Consequently, this approach produces
inputs that are syntactically valid by construction.  We analyzed
three grammar-based fuzzers--Grammarinator~\cite{grammarinator},
jsfunfuzz~\cite{jsfunfuzz}, and
QuickFuzz~\cite{quickfuzz,grieco2016quickfuzz}.  Unfortunately, none
of those was effective out of the box. Given the bound limits we
defined on the number of inputs to generate and the exploration depth
of the syntax tree, we were unable to find discrepancies on the inputs
those tools generated. To sum, the tools produced a high percentage of
semantically-invalid inputs (\eg{}, \Fix{give one or two concrete
  examples--variable usage without definition?}) that we needed to
discard and the valid inputs manifested no discrepancies. For example,
we produced 100K inputs with Grammarinator and only \Fix{X} of those
were valid. With QuickFuzz, we were able to produce \Fix{Y, Y$>$Y?}
valid inputs as it contains some heuristics to circumvent violation of
certain type rules such as \Fix{variable used must be
  defined?}. Nonetheless, running those inputs in our infrastructure
we were unable to find discrepancies. Inpecting those inputs, we
realized that they reflected very simple scenarios.\Comment{ \Mar{are there
  weights associated with production rules besides the depth you can
  set?}} Although grammar-based fuzzers has been shown effective to
find real bugs\Fix{cite some jsfunfuzz bug reports}, generating
valid inputs is very time-consuming, especially for the context of
differential testing. For this reason, we did not consider those fuzzers.

@@

Mutational fuzzers can be either white-box or black-box. White-box
mutational fuzzers are typically coverage-based. American Fuzz Loop
(AFL)~\cite{afl} and libFuzzer~\cite{libfuzzer} are examples of this
kind of fuzzers. These fuzzers run tests inputs against instrumented
versions of the program under testing with the typical goal of finding
univeral errors like crashes and buffer overflows. The instrumentation
adds code to collect branch coverage and to monitor specific
properties\footnote{There are options in the clang toolchain to build
  programs with fuzzing instrumentation~\cite{libfuzzer}. clang
  provides several sanitizers for property
  checking~\cite{clang-documentation}.}. AFL use coverage to determine
inputs that uncover a new branch and deserve fuzzing whereas libFuzzer
uses evolutionary generation--it tries to minimize the distances to
still-uncovered branches of the program. AFL takes the instrumented
program binary (say, a JS engine) and one seed input to that program
(say, a JS program) and produces on output fault-revealing inputs, if
found. Considering our context of application, we needed to instrument
one runtime engine for fuzzing; we chose v8. Unfortunately, we found
that most of the inputs produced by AFL violate the JS
grammar. Furthermore, the fuzzing task can take daysfor a single seed
input and there is no simple way to guide the
exploration~\Fix{cite}. That happens because the fuzzer aims to
explore the entire decision tree induced from the engine's main
function, including the branches associated with the lexer and the
parser. For those reasons, coverage-based fuzzers were not considered
in this study. It is worth mentioning that Google mitigates that
problem with libFuzzer by asking developers to create fuzz targets for
specific program
functions\cite{libFuzzer-tutorial-google,libFuzzer-chromium-google}. Although
that approach has shown very effective, it requires domain knowledge
to create the calling context to invoke the fuzz target.

\Fix{...do the same as above for black-box fuzzing. explain what kinds of
inputs they can generate.}

@@

>>>>

JavaScript (\js{}) is today one of the most popular programming
languages for the web~\cite{business-insider,stackify}. The interest
of the community for the language encourages constant improvements in
its specification and implementations (see
Section~\ref{sec:es6-design}). It is natural to expect that such
improvements will entail sensible changes in runtime
engines~\cite{kangax}; changes that could lead to errors. Even small
changes can have a large impact in practice. For example, in October
2014 a new attribute added to the Array datatype caused a bug that
resulted in MS Outlook web not working under
Chrome~\cite{array-bug-chromium-issue4247,array-bug-discussion}. To
sum, finding bugs in JS engines is an important problem specially
because of the speed the specification evolves and the availability of
various public engines.

Differential testing~\cite{Brumley-etal-ss07} (DT) is a popular
approach that has been applied in a variety of contexts to find bugs
in
software~\cite{Yang-etal-pldi11,Chen-etal-fse2015,Argyros-etla-ccs16,Chen-etal-pldi16,petsios-etal-sp2017,SivakornAPKJ17}. It
automates test generation in scenarios where multiple implementations
of a system exist. DT leverages the diversity across system's
implementations to detect anomalous behavior. For example, Mozilla
runs JS files--created with the grammar-based fuzzer
jsfunfuzz~\cite{jsfunfuzz}--against different configurations of their
SpiderMonkey engine~\cite{jsfunfuzz-mozilla-bug} to find discrepancies
across different configurations of the same build (\eg{}, enabling or
not eager JIT compilation). They use this technique since 2002 and
have been able to find over 270 bugs since
then~\cite{jsfunfuzz-at-mozilla}, including security bugs.
Cross-engine differential testing, in contrast, checks correctness by
comparing the output produced by \emph{different} engines on a given
input~\cite{patra2016learning}. Although different implementations of
JS engines today (see Table~\ref{tab:engines}), this technique is not
used in practice.

%% WHAT WE DID
This paper reports the results of a study we conducted to (i) assess
the reliability of existing JS engines and to (ii) evaluate the
importance of different factors when using DT. More precisely, we
analyzed the impact of two orthogonal factors in DT: test mining and
fuzzing. The first factor--\emph{test mining}--considers the effect
that \js{} test files mined from the repositories of open-source
engines have in detecting functional bugs in other engines. The second
factor--\emph{fuzzing}--considers the impact in bug detection of
fuzzing these files.
%% WHAT WE FOUND
To sum, we found that \Fix{...}

@@

\section{JavaScript}
\label{sec:es6-design}

JavaScript engines are virtual machines that parse source code and
compile it in bytecodes. These engines implement some version of
ECMAScript\footnote{The name JavaScript still prevails today,
  certainly for historical reasons.} (\es{}), which emerged with the
goal to standardize variants of the language, such as Netscape's
JavaScript and Microsoft's JScript. The \es{} specification is
regulated by Ecma International~\cite{es6-website} under the
TC39~\cite{tc39-github} technical committee.  Every year, a new
version of the \es{} specification is produced with new features and
minor fixes. The canonical spec today is
ES6~\cite{ecmas262-spec-repo,ecmas262-spec}. Features of these
versions and supporting engines are publicly available~\cite{kangax}.
In the following, we briefly describe different sources of
discrepancies that can emerge across different engine implementations
as results of the active changes in the \es{} specification.

%% \Mar{what is that exactly? can we cite? is that only spec or actually
%%   code as most of java.* libraries.}  \Igor{seems like an API. They
%%   define the language behavior; grammar; values to static variables;
%%   the methods for each "class"(Object)...  These irregularities
%%   appears on engine implementation, maybe an wrong if statement and so
%%   on.} \Mar{Igor, avoid general discussion like $\leftarrow$this. look for
%%   explanation in some forum/tutorial and explain **when** you
%%   undertand the JS ecosystem design and can explain with confidence.}

\subsection{Undefined (Implementation-Dependent) Behavior}
\label{sec:imp-dep-behavior}

The \es{} spec uses the label ``implementation-dependent'' to indicate
cases where each engine developer can decide how to implement some
aspect of a functionality. In these cases, execution may differ from
engine to engine. One of the reasons implementation-dependent behavior
exists is to enable compiler optimizations. For example, the
\mcodeid{for-in} loop does not clearly specify the iteration
order~\cite{so-forin-undefined,javascript-in-chrome} and different
engines capitalize on that~\cite{for-in-undefined}. As another
example, \Fix{...}

@@

\subsection{Randomness}

\sloppy Some elements in JS have non-deterministic behavior (\eg{},
Math.random and Date). A test that make decisions based on these
elements could, in principle, produce different outcomes on different
runs. Carefully-written test cases should not manifest this kind of
flakiness~\cite{luo-etal-fse2014,palomba-zaidman-icsme2017}, but
fuzzed test files are more likely to manifest them.

\subsection{\Fix{Other}}

1) The stack depth in JS is often close to the real (native) C stack
depth, nevertheless observable by JS through recursion. This causes
differential behavior even on the same engine when JITs are enabled or
disabled, as the stack frame sizes vary. LangFuzz in particular
produces a lot of recursions and hits these problems even more often
than jsfunfuzz.\Mar{ask C. Holler. it is unclear to me how the output
  could be different.}

\subsection{Bugs}

@@

@@

\subsection{Seed JS files\label{sec:seeds}}~
We selected test files from several public projects from GitHub. We
looked for self-contained tests from JS engine projects, initially
using the GitHub REST API~\cite{github-rest-api}. We noticed that some
of the tests we found depend on external libraries, which not all
engines we use (see Section~\ref{sec:methodology:engines}) support. We
decided to discard those. For example, many tests we found required a
Node.js runtime~\cite{node} for execution. The main source of test
cases is the offical TC-39~\cite{tc39-github} conformance test suite
for the ECMA262 specification~\cite{ecmas262-spec}. This test suite
contains 87\% of all the tests we used. Furthermore, we
considered test suites of three of the four engines mentioned in
Section~\ref{sec:methodology:engines} and four other engines.
Table~\ref{tab:test-suites} shows the JS sources we considered for
each of the selected projects. Overall, we found a total of
\totfiles{} JS files. It is worth noting that some test suites use
distinct frameworks for testing and require external libraries to run
the tests.  For these projects, we needed to make small
changes in the testing infrastructure to be able to run the tests
uniformly across all engines. More precisely, we needed to mock
some functions, related to the test harness, which are only available to
certain engines.

@@

Generational fuzzers are typically grammar-based. These fuzzers
generate a new file using the grammar of the language whose inputs
should be fuzzed. Intuitively, those fuzzers implement a traversal of
the production rules of the input grammar to create syntax trees,
which are then pretty-printed. Consequently, this approach produces
inputs that are syntactically valid by construction. We analyzed four
grammar-based fuzzers--Grammarinator~\cite{grammarinator},
jsfunfuzz~\cite{jsfunfuzz},
LangFuzz~\cite{Holler:2012:FCF:2362793.2362831}, and
Megadeth~\cite{grieco2016quickfuzz}.  Unfortunately, none of those
were effective out of the box. For example, we produced 100K inputs
with Grammarinator and only \Fix{X} of those were valid. With
Megadeth, we were able to produce \Fix{Y, Y$>$Y?}  valid inputs as it
contains some heuristics to circumvent violations of certain type
rules such as \Fix{variable used must be defined?}. Nonetheless,
running those inputs in our infrastructure we were unable to find
discrepancies. Inpecting those inputs, we realized that they reflected
very simple scenarios. To sum, a high percentage of inputs that
Grammarinator and Megadeth generated were semantically-invalid that we
needed to discard whereas the valid inputs manifested no
discrepancies. Considering jsfunfuzz~\cite{jsfunfuzz}, we noticed
that, in addition to the issues mentioned above, it produces inputs
that use functions that are only available in the SpiderMonkey
engine. We would need either to mock those functions in other engines
or to discard those tests. Considering
LangFuzz~\cite{Holler:2012:FCF:2362793.2362831}, the tool is not
publicly available. One fundamental issue associated with generational
fuzzers in our context is that the tests they produce do not contain
assertions; to enable the integration of this kind of fuzzers in our
infrastructure, we would need to look for discrepancies across
compiler messages.  All in all, although grammar-based fuzzers have
been shown effective to find real
bugs~\cite{Holler:2012:FCF:2362793.2362831}, generating valid inputs,
albeit promising conceptually, still requires some engineering
effort. For this reason, we did not consider those fuzzers in this
study.

@@

Mutational fuzzers can be either white-box or black-box. White-box
mutational fuzzers are typically coverage-based. American Fuzz Loop
(AFL)~\cite{afl} and libFuzzer~\cite{libfuzzer} are examples of this
kind of fuzzers. These fuzzers run tests inputs against instrumented
versions of the program under testing with the typical goal of finding
univeral errors like crashes and buffer overflows. The instrumentation
adds code to collect branch coverage and to monitor specific
properties\footnote{There are options in the clang toolchain to build
  programs with fuzzing instrumentation~\cite{libfuzzer}. clang
  provides several sanitizers for property
  checking~\cite{clang-documentation}.}. AFL use coverage to determine
inputs that uncover a new branch and hence should be fuzzed more
whereas libFuzzer uses evolutionary generation--it tries to minimize
the distances to still-uncovered branches of the program. AFL takes
the instrumented program binary (say, a JS engine) and one seed input
to that program (say, a JS program) and produces on output
fault-revealing inputs, if found. Considering our context of
application, we needed to instrument one runtime engine for fuzzing;
we chose v8. Unfortunately, we found that most of the inputs produced
by AFL violate the JS grammar. Furthermore, the fuzzing task can take
days for a single seed input and there is no simple way to guide the
exploration\footnote{Exchanged emails with the tool author.}. That
happens because the fuzzer aims to explore the entire decision tree
induced from the engine's main function, including the branches
associated with the higher layers of the compiler (\eg{}, lexer and
parser). It is worth mentioning that Google mitigates that problem
with libFuzzer by asking developers to create fuzz targets for
specific program
functions\cite{libFuzzer-tutorial-google,libFuzzer-chromium-google}. Although
that approach has shown very effective, it requires domain knowledge
to create the calling context to invoke the fuzz target. For those
reasons, coverage-based fuzzers were not considered in this study.

We uses two black-box mutational fuzzers in this
study--\radamsa~\cite{radamsa} and \quickfuzz~\cite{quickfuzz}. These
fuzzers require no instrumentation and domain-knowledge. They mutate
\emph{existing} inputs randomly. The strenght of the approach is
limited by the supported mutation operators, which are typically
simple, and the quality of the test suite. We chose \radamsa\ and
\quickfuzz\ \Fix{...}

@@

%%  LocalWords:  bytecodes JScript Ecma
